A high-repetition-rate, fast temperature-programmed gas chromatograph and its on-line coupling to a supercritical fluid chromatograph (SFC×GC)

### 2020-01-04 Journal/ResponseLetter.tex

% Title:    A LaTeX Template For Responses To a Referees' Reports
% Author:   Petr Zemek <s3rvac@gmail.com>
% Homepage: https://blog.petrzemek.net/2016/07/17/latex-template-for-responses-to-referees-reports/
% License:  CC BY 4.0 (https://creativecommons.org/licenses/by/4.0/)
\documentclass[10pt]{article}

% Allow Unicode input (alternatively, you can use XeLaTeX or LuaLaTeX)
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{eurosym}
\usepackage{textcomp}

\usepackage{microtype,xparse,tcolorbox}
\newenvironment{reviewer-comment }{}{}
\tcbuselibrary{skins}
\tcolorboxenvironment{reviewer-comment }{empty,
  left = 1em, top = 1ex, bottom = 1ex,
  borderline west = {2pt} {0pt} {black!20},
}
\ExplSyntaxOn
\NewDocumentEnvironment {response} { +m O{black!20} } {
  \IfValueT {#1} {
    \begin{reviewer-comment~}
      \setlength\parindent{2em}
      \noindent
      \ttfamily #1
    \end{reviewer-comment~}
  }
  \par\noindent\ignorespaces
} { \bigskip\par }

\NewDocumentCommand \Reviewer { m } {
  \section*{Comments~by~Reviewer~#1}
}
\ExplSyntaxOff
\AtBeginDocument{\maketitle\thispagestyle{empty}\noindent}

% You can get probably get rid of these definitions:
\newcommand\meta[1]{$\langle\hbox{#1}\rangle$}
\newcommand\PaperTitle[1]{``\textit{#1}''}

\title{Response letter of \meta{RSI19-AR-01677} \\
  Based on the Referees' Report}
\author{D. Malan \and S.J. van der Walt \and E.R. Rohwer}
\date{\today}

\begin{document}

\noindent
\textbf{Chin-Tu Chen}\\
\textbf{Associate Editor}\\
\textbf{Review of Scientific Instruments}

\medskip

\begin{quote}
Revision of the paper \meta{RSI19-AR-01677} submitted to the Review of
Scientific Instruments, entitled \PaperTitle{A high-repetition-rate, fast
temperature programmed gas chromatograph and its on-line coupling to a
supercritical fluid chromatograph (SFC×GC)}, based on the referees' report.
\end{quote}

\medskip

\noindent
Dear Chin-Tu Chen:

Thank you for your email dated 7 November 2019 enclosing the reviewers’ comments. We
have carefully reviewed the comments and have revised the manuscript accordingly. Our
responses are given in a point-by-point manner below.

We hope the revised version is now suitable for publication and look forward to hearing from you
in due course.

\vspace{\baselineskip}

\noindent
Yours sincerely,\\
D. Malan\\
Department of Chemistry\\
University of Pretoria

\Reviewer{\#1}
\begin{response}{Abstract, last line: the sentence fragment 'separating fatty information" needs revision.
	}
 	The sentence was revised and clarified.
\end{response}

\begin{response}{p3, orthogonal and general elution problem. In trying to wrap
my head around this argument, I can't help but think that despite their comment
about GC×GC wrap-around, it does seem that this would be even more of a problem
in their current setup or those using LC-GC. The authors miss the opportunity to
justify the need for fast temperature programmed GC in their second dimension.
Perhaps laying their arguments out in more detail here would be useful.
} 
The problem with the vagueness of the argument was identified. The paragraphs
were changed to clarify how the general elution problem applies to
multidimensional chromatography, and to emphasize that SFC×GC demands
temperature programming in the GC dimension.
\end{response}

\begin{response}{p6 last line, sentence structure needs revision.
	}  The sentence was revised and clarified.
\end{response}

\begin{response}{p7, description of the connection blocks. Refer the reader to see Figure 1 here. 
	}  A reference to Figure 1 was added.
\end{response}

\begin{response}{Also, singular-plural confusion in :"These blocks fulfilled three roles: It (1) sealed ... "
	}  
	The confusion was cleared up. 
\end{response}

\begin{response}{Figure 1. It is not clear from the drawing that the cryogenic
coolant actually enters the heater tube. While the dimensions of the drawing are
perhaps selected to be according to scale, the reader would benefit from a
little artistic liberty here. (g) "electrical connection" is a little vague.
	} The figure was changed to add a schematic diagram that illustrate the coolant
	flow, and the caption was clarified.
\end{response}

\begin{response}{p9. Reference resistor and ballast resistor. What are the
dimensions of the wires used to construct these resistors? What is the
resistance of the ballast? 
	} 
	The requested details were added, as well as a figure that shows the construction of the relevant resistors.
\end{response}

\begin{response}{p13, A. Heating uniformity. Text says video is shown in
Figure3. Figure 3 is a screen capture of the video shown in Supplemental
Material. Correct accordingly.
	}
	We recognize the inaccuracy, and the text was corrected. 
\end{response}

\begin{response}{ p13. Gradients by cooling. Not sure how this statement
accurately reflects the repeated heating cycles shown in Figure 4. Further
information and a more detailed description of Figure 4 is required. Does the
difference between the two calibration points along the length of the column (33
and 66 cm) increase with temperature? 
	}  
	The discussion on the gradients measured by thermocouple was revised and expanded for greater clarity. 
\end{response}

\begin{response}{Figure 5. Explain in the heading the blue and red lines.
	}  
	The caption was revised with a better description that includes an explanation of the colours.
\end{response}

\begin{response}{Figure 2. Can be improved by showing more details of the
circuit, rather than just three resistors in series.
	}  The figure was replaced by a more detailed one, and the description in the text was adapted to match the figure.   
\end{response}

\begin{response}{A few references appear incomplete. e.g. 17 and 29.}  
	The bibliography was reviewed and all the incomplete references were updated.
\end{response}

\Reviewer{\#2}

\begin{response}{p1 Abstract, sentence change: We demonstrate the fast
chromatograph by separating fatty acid methyl esters by carbon number, which
would be useful in the food and biodiesel industries.
	}  
	The sentence was clarified. 
\end{response}

\begin{response}{p2: This concept is called comprehensively coupled
chromatography. For example, in comprehensively coupled gas chromatography
(GC×GC). Confirm correct nomenclature in as published by Marriot in LCGC 2012.
	} We thank the reviewers for drawing attention to using correct nomenclature. We
	revised the manuscript to ensure the recommended nomenclature was used throughout.
\end{response}

\begin{response}{p2 Sentence change: The use of narrow-bore columns and thin
films reduce run times by improving mass transfer between the mobile and
stationary phases, resulting in analytes spending more time in the mobile phase
allowing shorter retention times.

	} We agree that the sentence needed clarification, but disagree with the
	explanation. Improved mass transfer does not allow analytes to spend more time
	in the mobile phase, but it does reduce peak broadening.  The sentence was
	revised thusly: \textit{Typically, short columns and high carrier flows rates
	are used, which reduce retention times by reducing void times. Resolution can be
	maintained at even higher gas flows by using narrow-bore columns and thin films:
	these improve the mass transfer rate between the mobile and stationary phases,
	which reduces peak broadening by bringing the analyte exchange between the
	phases closer to equilibrium at higher-than-optimum carrier gas velocities.}

\end{response}

\begin{response}{p3 sentence change: it becomes unlikely that one set of
acceptable operating conditions (temperature, flow and stationary phase) will
give acceptable (fast enough and with adequate resolution) separation in both
dimensions.
	}  This sentence was revised as part of general improvement in the argument.
	(See Reviewer 1, comment ``p.3'')
\end{response}

\begin{response}{p3 sentence change: so late that the next modulation period has
already started before its elution and the peak only appears on the next 2D
chromatogram 
	}  The sentence was revised as suggested.
\end{response}

\begin{response}{p3 sentence change: If wraparound is a problem in GC×GC, it
becomes inevitable in SFC×GC since the first dimension (SFC) does not normally
separate compounds in a way that correlates with vapour pressure.
	}  This sentence was improved by revising the paragraph. 
\end{response}

\begin{response}{p5 sentence change: The Zip ScienticTM solution operates by forced convection.}  
	The sentence was changed to the suggested one. 
\end{response}

\begin{response}{p5: sentence change: In our case we applied it as a fast second
dimension separation, comprehensively coupled to a supercritical fluid
chromatograph.
	}  The sentence was changed to the suggested one. 
\end{response}

\begin{response}{p6 clarification: clarify what type of static restriction was
used.
	}  The description of the static restrictor was improved. 
\end{response}

\begin{response}{p6: sentence change: the stop valve opens and the SFC eluate
exits the restrictor into the hot splitless inlet of the GC, where it evaporates
and is transported onto the GC column.
	}  This sentence was improved by rewriting the paragraph. 
\end{response}

\begin{response}{p6: sentence change: which of course means the start of the
temperature program, which desorbs the analyteS trapped on the head of the cold
column.
	}  This sentence was improved by rewriting the paragraph.
\end{response}

\begin{response}{p7: sentence change: The split vent valve was disconnected from
the gas chromatograph CONTROL and CONNECTED TO the control computer described
below 
	}  The sentence was changed to the suggested one. 
\end{response}

\begin{response}{p10: Figure 2: Figure do not contain that much detail and may
be omitted.
	}  The figure was replace by a more detailed one, and the text adapted to match
	the figure. (See Reviewer 1, comment ``Figure 2'')
\end{response}

\begin{response}{p12: abbreviation: LabVIEW 7.1TM (National Instruments). This
abbreviation was used earlier p10 and should be clarified there.
	}  We recognize the oversight and corrected it.
\end{response}

\begin{response}{p15: sentence change: In the instrument presented here, THE
power was limited by the voltage output range of the DAC of the data acquisition
board, and not by the power of the heater.
	}  The sentence was replaced by the suggested one.
\end{response}

\end{document}

### 2020-01-04 document.tex
Final version as submitted after adding response to reviewers.

\documentclass[aip,rsi,preprint,graphicx]{revtex4-1} % for checking your page length
%\documentclass[aip,rsi,preprint,graphicx]{revtex4-1} % for review purposes
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{siunitx}

@@

We present a fast gas chromatographic system that can be used as a second
dimension in comprehensive (supercritical fluid × gas) chromatography (SFC×GC).
The temperature of the short (1 metre long) capillary column is controlled by a
resistively heated coaxial stainless-steel tube. The electrical resistance and
therefore temperature of the stainless-steel tube is measured by continuous
monitoring of the current/voltage ratio. Highly repeatable heating rates of up
to \SI{2100}{\celsius\per\minute} (\SI{35}{\celsius\per\second}) are obtained,
which should be high enough for the most demanding fast chromatograms. To reduce
the cooling time between temperature programs the column is cooled by injecting
evaporating carbon dioxide into the space between the coaxial heater and the
column. This gives cooling rates of \SI{5100}{\celsius\per\minute}
(\SI{85}{\celsius\per\minute}) which allows quick succession of temperature
programs. More repeatable heating profiles with stable GC retention times
together with faster cooling are significant improvements on previous SFC×GC
systems. Cycle times of four gas chromatograms per minute could readily be
achieved, which allows efficient coupling to high-resolution stop-flow SFC in
the first dimension. We demonstrate the fast chromatograph by separating fatty
information which would be useful in the food and biodiesel industries.

@@

\section{Introduction}
\label{Introduction}
% Niel: Literature review
Speed of analysis is a most important criterion for any analytical system.
Chromatography as an analytical technique is notoriously slow, compared to, say,
some spectroscopic techniques.

While it is of interest to (especially high-throughput) laboratories to reduce
run times, at some point the faster chromatography starts outstripping the
sample preparation capacity. Once the chromatographic run duration approaches
the time it takes to prepare the sample, faster chromatography will not improve
the sample throughput. Technology development in fast chromatography is
therefore driven not by conventional injection-based chromatography, but by
applications where no sample preparation is necessary and rapid results are
needed.

@@

However, in the past two decades the major application for fast gas
chromatography has not been in direct sample analysis, but in the analysis of
fractions of eluate from chromatographic separations. This concept is called
\textit{comprehensively coupled chromatography}. For example, in comprehensively
coupled gas chromatography (GC×GC), fractions of a conventional gas
chromatographic run are collected by a \textit{modulator} and then admitted to a
second, fast gas chromatograph\cite{Liu1991}. GC×GC has become well established:
the annual GC×GC Symposium is in its 15th year, and major reviews are published
regularly\cite*{Seeley2013,Prebihalo2018}.

In comprehensive 2D chromatography the second-dimension chromatography must be
\textit{fast}. The theory and practice of improving the speed of GC is well
developed\cite*{Cramers1999,Korytar2002}. Typically, short columns and high
carrier flows rates are used, which reduce retention times by reducing the
void time. The use of narrow-bore columns and thin films reduce run times by
improving mass transfer between the mobile and stationary phases, allowing
higher linear flow rates of the carrier gas.

Comprehensively coupled chromatography depends on \textit{orthogonality}: the
second dimension ($^2$D) separation must have a different mechanism than that of
the first dimension ($^1$D) separation\cite*{Giddings1995,Camenzuli2014}. The
higher the orthogonality, the better the \textit{separation space} offered by
the system. The different separation mechanisms that determine orthogonality are
usually provided by the different stationary phases offered by column
manufacturers, but it is also possible to use a different form of chromatography
as a first dimension: liquid chromatography has been used as a first dimension
in LC×GC\cite{Koning2004}, and in our laboratory we are developing
SFC×GC instrumentation\cite{Venter2004}.

If a chromatographic $^2$D separation is highly orthogonal to the $^1$D
separation, then we will run into the \textit{general elution
problem}\cite{Skoog2007}: it becomes unlikely that one set of acceptable
operating conditions (temperature, flow and stationary phase) will give
acceptable (fast enough and with adequate resolution) separation.

In GC×GC as it is practised today, this has become a real problem.
\textit{Wraparound} is a well-known phenomenon in
GC×GC\cite{Dalluege2003}. Wraparound happens when a peak or peaks elute
`late' on the second dimension, so late that the next modulation period has
already started before its elution and the peak appears on the next $^2$D
chromatogram.

% There are ways to avoid or ignore wraparound, but using gradient elution in
% the second dimension would, in principle, prevent all wraparound.

If wraparound is a problem in GC×GC, it becomes inevitable in
SFC×GC. The first dimension (SFC) does not normally separate compounds in
a way that correlates with vapour pressure. Indeed, SFC is particularly good in
performing group-type separations\cite{Venter1999}, where compounds with a wide
range of boiling points elute together.

\subsection{Column heating for gas chromatography}

In GC the usual solution to the general elution problem is temperature
programming, \textit{i.e.} increasing the temperature of the column throughout
the duration of the chromatograpic run. Temperature programming decreases the
retention times of late-eluting analytes without sacrificing the resolution of
early-eluting analytes. Instrument manufacturers have mastered the art of
temperature programming, and today the vast majority of 1D gas chromatography
uses temperature programming.

@@

\subsection{Cooling of the GC column}

@@

There have been attempts to improve the cooling rate of air baths.

Agilent\texttrademark{} markets a `low thermal mass' column, which includes a
collinear heating wire and a collinear sensing element bundled with a short
silica column. These units cool down faster than normal air-baths, but only to
ambient temperature. For short columns a cooling time of \SI{1.3}{\minute} has
been reported\cite{Luong2006}. These columns are available with only a limited
range of stationary phases.

The Zip Scientific\texttrademark{} solution is forced convection. Their
`GC-Chaser' system uses a blast of (chilled) room air to the
GC oven to cool it down after each GC run, allowing the use of conventional
fused silica columns. In an example\cite{ZipScientific2008} it reduces cooling
time from 16 minutes to 7 minutes. Although this can significantly decrease the
cost per analysis in a high-throughput laboratory, the cooling is not fast
enough for 2D chromatography.

The EZ Flash\texttrademark{} system uses a coaxial resistive heating system,
which allows the column to cool down ballistically. Reports from the literature
claims cool-down periods of 30 seconds\cite{Dalluege1999}.

The Valco FTP-200 fast temperature programmer\cite{VICIAGI2019} (Valco
Instruments Co. Inc.) uses a small, optional fan to cool down the columns.

Cryogenic coolers for GC ovens are available but are intended to cool the oven
to low temperatures for the analysis of volatile compounds and not necessarily
to improve cycle times. In previous SFC×GC work in our
laboratories\cite{Venter2004, Venter2006} the cryogenic function of a Varian
3300 gas chromatograph was used to cool down the column oven at the beginning of
each run, using large quantities of coolant in the process. The resistively
heated column cooling required 30 seconds to get back to the starting
temperature in this cool environment. We estimate that a single SFC×GC run
consumed an unacceptable 15 kg of carbon dioxide coolant.

@@

The fast GC system described in this paper can be used in any application where
fast chromatograms are needed in rapid succession, for example in process
control. In our case we applied it as a $^2$D column comprehensively coupled to
a supercritical fluid chromatograph.

We applied our fast GC system to the evaluation of fatty acid profiles of
potential biodiesel feedstock. The fast GC separation was highly orthogonal to
the $^1$D separation on SFC, showcasing this type of comprehensively coupled
chromatography. An easily interpretable 2D separation was obtained, with run
times similar to the total run times required by official methods for the
determination of fatty acid profiles.

@@

For better understanding of subsequent detail, a description of the cycle of the
SFC×GC is appropriate.

The SFC subsystem runs in stop-flow mode. The eluate of the SFC passes through a
stop valve and a static capillary restrictor. The end of the restrictor is 
inserted through the septum into the hot inlet of the GC.

After the sample is injected, the SFC runs for an optional uninterrupted period,
usually briefer than the void time. No data is collected during this time. Then
the stop valve closes, and the cooling system activates to cool the GC column.
Once the column is at the set temperature, the stop valve opens and the SFC
eluate exits the restrictor into the hot splitless inlet of the GC, where it
evaporates and is transported on to the column. The vapor-phase material
is trapped in the cold stationary phase. Once the desired amount
of the SFC eluate has been collected, the stop valve closes. Some time is
allowed to pass so that the last of the evaporated SFC eluate flushes into the
column. Then the split vent of the GC inlet opens for a short period, which
vents excess carbon dioxide to the atmosphere. This normalizes the pressure in
the inlet (having been raised  by the high flow rate of CO$_2$ from the SFC) and
allows the carbon dioxide to be replaced by the pressure-controlled hydrogen
carrier gas. Once the split vent closes the GC run starts, which of course means
the start of the temperature program, which desorbs the analyte trapped on the
head of the cold column. At the end of the GC run, the cooling process starts,
cooling the column in a second or two. The next cycle then starts, marked by the
opening of SFC the stop valve.

\subsection{Hardware}
The short-cycle fast gas chromatograph was built into a modified
Varian\texttrademark{} 3300 gas chromatograph.

The inlet was an unmodified Varian\texttrademark{} 1075 split/splitless inlet.
The temperature control of the inlet was by the original electronics, controlled
from the Varian 3300 front panel. The split vent valve was disconnected from the
gas chromatograph and controlled from the control computer described below.

The detector was an unmodified Varian\texttrademark{} 3300 flame ionization
detector (FID). The temperature of the detector was controlled by the original
electronics, controlled from the front panel. The detector bias voltage was
supplied by the original electronics, but a stand-alone high-speed electrometer
(V.G. Micromass Ltd, Model M406-H) captured the signal, which was then
conditioned by a bench-top amplifier (V.G. Micromass Ltd, Model M406) before it
was sent to the computer.

The coaxial heater was made of a 940 mm length of stainless steel (SAE 304
grade), obtained from MIFAM (Milanówek, Poland). It had an outside diameter of
1.06 mm and an inside diameter of 0.80 mm.

The ends of the coaxial heater tube terminated in two identical, specially
designed T-piece blocks. These blocks fulfilled three roles: It (1) sealed the
coaxial tube to contain coolant, (2) allowed electrical connection to the
coaxial heater, and (3) acted as a heated transfer line between the column and
the detector and inlet.

The T-piece blocks were machined from brass. The stainless-steel tube of the
coaxial heater was brazed onto the block. The cooling CO$_2$ was let in from one
side, and on the opposite side a thermocouple probe inserted into a blind hole
measured the temperature of the block. A micro-union using a metal ferrule
(Restek\texttrademark{} SilTite $\mu$-Union Connectors), brazed to the block,
was used to seal the capillary column. An electrical contact was added, to which
an electrical conductor could be soft-soldered. Four holes through the length of
the block allowed the fitting of four 220V, 100 W heater cartridges (Hotset),
providing each of the two blocks 400W of heating.

\begin{figure}
	\includegraphics[width=\textwidth]{./Figures/Cutaway_with_symbols.png}
	\caption["The T-piece block"]{A cutaway diagram of the inlet/outlet outlet T-piece blocks, showing the (a) cryogenic coolant inlet/outlet, (b) the coaxial heater, (c) the capillary column, (d) the mico-union seal, (e) the cartridge heater socket, (f) the thermocouple socket, and (g) the electrical connection.}%
	\label{fig:Manifold}
\end{figure}

The chromatographic column entered the T-piece block via the bottom port with the
coaxial heater tube and passed out through the top port and into the inlet or
detector. In the left-hand side of the T-piece block was the inlet or outlet port for the
carbon dioxide coolant.

%A blind hole on the right-hand side of the manifold mounted a thermocouple. The
%temperature reported by the thermocouple was fed to control software, which
%controlled the temperature of the manifold.

Since the coaxial heater resistance measurement used a two-wire technique, the
current-carrying circuit was made entirely of soldered joints to minimize stray
resistance from developing in electrical connections.

@@

The cryogenic carbon dioxide (Afrox, Tec (Wet)) was supplied from a cylinder
equipped with a dip tube. It flowed to a coil heat exchanger bathed in a cold
\SI{-5}{\celsius}) propylene glycol heat transfer fluid, which was mounted on
top of the Varian GC. From the heat exchanger the carbon dioxide flowed down to
the cryo shut-off valve (ASCO, RedHat brand). This arrangement ensured a supply
of liquid CO$_2$ at the shut-off valve, without which some gaseous CO$_2$ might
have to vent through the coaxial heater before the liquid CO$_2$ coolant could
reach it. Downstream of the shut-of valve there was a 10-turn metering valve,
which controlled the flow of carbon dioxide into the coaxial heater for
evaporative cooling. At the outlet end of the coaxial heater the carbon dioxide
escaped through the T-piece block into the atmosphere.

>>>>

\documentclass[aip,rsi,preprint]{revtex4-2} % for checking your page length
%\documentclass[aip,rsi,preprint,graphicx]{revtex4-1} % for review purposes
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{siunitx}

@@

We present a fast gas chromatographic system that can be used as a second
dimension in comprehensive two-dimensional (supercritical fluid × gas)
chromatography (SFC×GC). The temperature of the short (1 metre long) capillary
column is controlled by a resistively heated coaxial stainless-steel tube. The
electrical resistance and therefore temperature of the stainless-steel tube is
measured by continuous monitoring of the current/voltage ratio. Highly
repeatable heating rates of up to \SI{2100}{\celsius\per\minute}
(\SI{35}{\celsius\per\second}) are obtained, which should be high enough for the
most demanding fast chromatograms. To reduce the cooling time between
temperature programs the column is cooled by injecting evaporating carbon
dioxide into the space between the coaxial heater and the column. This gives
cooling rates of \SI{5100}{\celsius\per\minute} (\SI{85}{\celsius\per\second})
which allows quick succession of temperature programs. More repeatable heating
profiles with stable GC retention times together with faster cooling are
significant improvements on previous SFC×GC systems. Cycle times of four gas
chromatograms per minute could readily be achieved, which allows efficient
coupling to high-resolution stop-flow SFC in the first dimension. We demonstrate
the fast chromatograph by separating fatty acid methyl esters, yielding
information which would be useful in the food and biodiesel industries.

@@

\section{Introduction}
\label{Introduction}
% Niel: Literature review
Speed of analysis is a most important criterion for any analytical system.
Chromatography as an analytical technique is notoriously slow, compared to, say,
some spectroscopic techniques. While it is of interest to (especially
high-throughput) laboratories to reduce run times, at some point the faster
chromatography starts outstripping the sample preparation capacity. Once the
chromatographic run duration approaches the time it takes to prepare the sample,
faster chromatography will not improve the sample throughput. Technology
development in fast chromatography is therefore driven not by conventional
injection-based chromatography, but by applications where no sample preparation
is necessary and rapid results are needed.

@@

However, in the past two decades the major application for fast gas
chromatography has not been in direct sample analysis, but in the analysis of
fractions of eluate from chromatographic separations. This concept is called
\textit{comprehensive two-dimensional chromatography}. For example, in comprehensive
two-dimensional gas chromatography (GC×GC), fractions of a conventional gas
chromatographic run are collected by a \textit{modulator} and then admitted to a
second, fast gas chromatograph\cite{Liu1991}. GC×GC has become well established:
the annual GC×GC Symposium is in its 15th year, and major reviews are published
regularly\cite*{Seeley2013,Prebihalo2018}.

In comprehensive 2D chromatography the second-dimension chromatography must be
\textit{fast}. The theory and practice of improving the speed of GC is well
developed\cite*{Cramers1999,Korytar2002}. Typically, short columns and high
carrier flows rates are used, which reduce retention times by reducing void
times. Resolution can be maintained at even higher gas flows by using
narrow-bore columns and thin films: these improve the mass transfer rate between
the mobile and stationary phases, which reduces peak broadening by bringing the
analyte exchange between the phases closer to equilibrium at higher-than-optimum
carrier gas velocities.

Comprehensive two-dimensional chromatography depends on \textit{orthogonality}:
the second dimension ($^2$D) separation must have a different mechanism than
that of the first dimension ($^1$D)
separation\cite*{Giddings1995,Camenzuli2014}. The higher the orthogonality, the
better the \textit{separation space} offered by the system. In GC×GC the
different separation mechanisms that determine orthogonality are usually
provided by the different stationary phases offered by column manufacturers, but
it is also possible to use a different form of chromatography as a first
dimension: liquid chromatography has been used as a first dimension in
LC×GC\cite{Koning2004}, and in our laboratory we are developing SFC×GC
instrumentation\cite{Venter2004}.

If a chromatographic $^2$D separation is highly orthogonal to the $^1$D
separation, then the $^2$D separation will run into the \textit{general elution
problem}\cite{Skoog2007}: the higher the orthogonality, the more diverse the
compounds found in each $^1$D fraction, and it becomes increasingly unlikely
that one set of acceptable operating conditions (temperature, flow and
stationary phase) will an give acceptable (fast enough and with adequate
resolution) $^2$D separation. In GC×GC as it is practised today this has become
a real problem, known as \textit{wraparound}, a well-known
phenomenon\cite{Dalluege2003}. Wraparound happens when a peak or peaks elute
`late' on the second dimension, so late that the next modulation period has
already started before its elution and the peak only appears on the next $^2$D
chromatogram. Careful interpretation of 2D chromatograms can identify which
peaks can be attributed to wraparound, and adjusting some parameters might
mitigate the problem.

If the general elution problem causes the phenomenon of wraparound in GC×GC,
where the compounds found in $^1$D fractions have a relatively narrow boiling
point range, it will cause much greater problems in SFC×GC, since the first
dimension (SFC) does not normally separate compounds in a way that correlates
with vapour pressure. Indeed, SFC is particularly good in performing group-type
separations\cite{Venter1999}, where compounds with a wide range of boiling
points elute together. Therefore, in SFC×GC we can expect compounds with a wide
range of boiling points in each $^1$D (SFC) fraction, and the general elution
problem makes it unlikely that an isothermal $^2$D (GC) run will provide an
acceptable separation. This is why successful SFC×GC demands temperature
programming in the GC dimension.

\subsection{Fast column heating for GC}

In GC the usual solution to the general elution problem is temperature
programming, \textit{i.e.} increasing the temperature of the column throughout
the duration of the chromatographic run. Temperature programming decreases the
retention times of late-eluting analytes without sacrificing the resolution of
early-eluting analytes. Instrument manufacturers have mastered the art of
temperature programming, and today the vast majority of 1D gas chromatography
uses temperature programming.

@@

\subsection{Fast column cooling for GC}

@@

There have been attempts to improve the cooling rate of air baths.
Agilent\texttrademark{} markets a `low thermal mass' column, which includes a
collinear heating wire and a collinear sensing element bundled with a short
silica column. These units cool down faster than normal air-baths, but only to
ambient temperature. For short columns a cooling time of \SI{1.3}{\minute} has
been reported\cite{Luong2006}. These columns are available with only a limited
range of stationary phases. The Zip Scientific\texttrademark{} solution operates
by forced convection. Their `GC-Chaser' system uses a blast of (chilled) room
air to the GC oven to cool it down after each GC run, allowing the use of
conventional fused silica columns. In an example\cite{ZipScientific2008} it
reduces cooling time from \SI{16}{\minute} to \SI{7}{\minute}. Although this can
significantly decrease the cost per analysis in a high-throughput laboratory,
the cooling is not fast enough for 2D chromatography. The EZ
Flash\texttrademark{} system uses a coaxial resistive heating system, which
allows the column to cool down ballistically. Reports from the literature claims
cool-down periods of \SI{30}{\second} \cite{Dalluege1999}. The Valco FTP-200
fast temperature programmer\cite{VICIAGI2019} (Valco Instruments Co. Inc.) uses
a small, optional fan to cool down the columns.

Cryogenic coolers for GC ovens are available but are intended to cool the oven
to low temperatures for the analysis of volatile compounds and not necessarily
to improve cycle times. In previous SFC×GC work in our
laboratories\cite{Venter2004, Venter2006} the cryogenic function of a Varian
3300 gas chromatograph was used to cool down the column oven at the beginning of
each run, using large quantities of coolant in the process. The resistively
heated column cooling required \SI{30}{\second} to get back to the starting
temperature in this cool environment. We estimate that a single SFC×GC run
consumed an unacceptable \SI{15}{\kilo\gram} of carbon dioxide coolant.

@@

The fast GC system described in this paper can be used in any application where
fast chromatograms are needed in rapid succession, for example in process
control. In our case we applied it as a fast second dimension separation,
comprehensively coupled to a supercritical fluid chromatograph.

We applied our fast GC system to the evaluation of fatty acid profiles of
potential biodiesel feedstock. The fast GC separation was highly orthogonal to
the $^1$D separation on SFC, showcasing this type of comprehensive
two-dimensional chromatography. An easily interpretable 2D separation was
obtained, with run times similar to the total run times required by official GC
methods for the determination of fatty acid profiles.

@@

For better understanding of subsequent detail, a description of the cycle of the
SFC×GC is appropriate. The SFC subsystem runs in stop-flow mode. The eluate of
the SFC passes through a stop valve and a static linear capillary restrictor.
The end of the restrictor is inserted through the septum into the hot inlet of
the GC. After the sample is injected, the SFC runs for an optional uninterrupted
period, usually briefer than the void time. No data is collected during this
time. Then the stop valve closes, and the cooling system activates to cool the
GC column. Once the column is at the set temperature, the stop valve opens and
the SFC eluate exits the restrictor into the hot splitless inlet of the GC,
where it evaporates instantly. The vapour-phase material is swept from the inlet
onto the column where it is trapped in the cold stationary phase. Once the
desired amount of the SFC eluate has been collected, the stop valve closes and
some time is allowed to elapse so that the last of the evaporated SFC eluate can
be flushed into the column. Then the split vent of the GC inlet opens for a
short period to vent excess carbon dioxide to the atmosphere. This normalizes
the pressure in the inlet (having been raised  by the high flow rate of CO$_2$
from the SFC) and allows the carbon dioxide to be replaced by the
pressure-controlled hydrogen carrier gas. The closing of the split vent marks
the start of the GC run, which of course means the start of the temperature
program. Under control of the temperature program the analytes trapped on the
head of the cold column desorb, migrate, separate, elute, and
are detected. The end of the temperature program marks the end of the GC run,
and the GC column is cooled to prepare for trapping the next fraction.

\subsection{Hardware}
The short-cycle fast gas chromatograph was built into a modified
Varian\texttrademark{} 3300 gas chromatograph. The inlet was a
Varian\texttrademark{} 1075 split/splitless inlet. The temperature control of
the inlet was by the original electronics, controlled from the Varian 3300 front
panel. The split vent valve was disconnected from the gas chromatograph control
and connected to the control computer described below.

The detector was an unmodified Varian\texttrademark{} 3300 flame ionization
detector (FID). The temperature of the detector was regulated by the original
electronics, controlled from the front panel. The detector bias voltage was
supplied by the original electronics, but a stand-alone high-speed electrometer
(V.G. Micromass Ltd, Model M406-H) captured the signal, which was then
conditioned by a bench-top amplifier (V.G. Micromass Ltd, Model M406) before it
was sent to the computer.

The coaxial heater was made of a \SI{940}{\milli\metre} length of stainless
steel (SAE 304 grade), obtained from MIFAM (Milanówek, Poland). It had an
outside diameter of \SI{1.06}{\milli\metre} and an inside diameter of
\SI{0.80}{\milli\metre}. The ends of the coaxial heater tube terminated in two
identical, specially designed T-piece blocks, machined from brass (See Figure
\ref{fig:Manifold}). These blocks fulfilled three roles:
They (1) sealed the coaxial tube to contain coolant, (2) allowed electrical
connection to the coaxial heater, and (3) acted as a heated transfer line
between the column and the detector and inlet. Each end of the stainless-steel
tube of the coaxial heater was brazed onto a block. The CO$_2$ coolant passed
through one side, and on the opposite side a thermocouple probe inserted into a
blind hole measured the temperature of the block. An electrical contact was
added, to which an electrical conductor could be soft-soldered. Four holes
through the width of the block allowed the fitting of four \SI{220}{\volt},
\SI{100}{\watt} heater cartridges (Hotset), providing each of the two blocks
\SI{400}{\watt} of heating.  The chromatographic capillary column entered the
T-piece block via the bottom port with the coaxial heater tube and passed out
through the top port (where a micro-union using a metal ferrule
(Restek\texttrademark{} SilTite $\mu$-Union Connectors) brazed to the block
created a seal to contain the coolant) and into the inlet or detector.

\begin{figure}
	\includegraphics[width=\textwidth]{T-Piece.eps}

\caption["The T-piece block"]{Left: A cutaway diagram of the inlet/outlet outlet
T-piece blocks with all the parts to scale, showing (a) the cryogenic coolant
inlet/outlet, (b) the coaxial heater, (c) the capillary column, (d) the
micro-union seal, (e) the cartridge heater socket, (f) the thermocouple socket,
and (g) the connector to which the power feed conductor was soft-soldered.
Right: A schematic diagram showing the flow of the coolant through the T-piece
block into the coaxial heater.}%

	\label{fig:Manifold}
\end{figure}

@@

The cryogenic carbon dioxide (Afrox, Tec (Wet)) was supplied from a cylinder
equipped with a dip tube. It flowed to a coil heat exchanger mounted on top of
the Varian GC from where the carbon dioxide flowed down to the cryo shut-off
valve (ASCO, RedHat brand). (The heat exchanger was bathed in a cold
(\SI{-5}{\celsius}) propylene glycol heat transfer fluid.) This arrangement
ensured a supply of liquid CO$_2$ at the shut-off valve, without which some
gaseous CO$_2$ might have to vent through the coaxial heater before the liquid
CO$_2$ coolant could reach it. Downstream of the shut-of valve a 10-turn
metering valve controlled the flow of carbon dioxide into the coaxial heater. At
the outlet end of the coaxial heater the carbon dioxide escaped through the
T-piece block into the atmosphere.

The SFC subsystem consisted of five packed silica columns
(\SI{150}{\milli\metre} × \SI{4.6}{\milli\metre}, \SI{3}{\micro\metre}
particles) (Restek, Pinnacle DB Silica) in series, a sample valve (VICI), and a
stop valve (VICI). Carbon dioxide, (\SI{99.995}{\percent} purity, Air Products)
was supplied to a Varian 8500 piston pump. The inlet pressure of the SFC columns
was controlled by a software proportional controller driving the stepper motor
of the piston pump. The high pressure at the column exit was maintained by a
linear fused silica capillary restrictor with an internal diameter of
\SI{0.050}{\milli\metre}, \SI{800}{\milli\metre} long.
